---
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(tidyquant)
```


# Youth Risk Behavior Surveillance

Every two years, the Centers for Disease Control and Prevention conduct the [Youth Risk Behavior Surveillance System (YRBSS)](https://www.cdc.gov/healthyyouth/data/yrbs/index.htm) survey, where it takes data from high schoolers (9th through 12th grade), to analyze health patterns.

## Load the data

This data is part of the `openintro` textbook and we can load and inspect it. There are observations on 13 different variables, some categorical and some numerical. 

```{r}
data(yrbss)
glimpse(yrbss)
```

## Exploratory Data Analysis

We first start with analyzing the `weight` of participants in kilograms. Using visualization and summary statistics, describe the distribution of weights. How many observations are we missing weights from?

```{r, eda_on_weight}

skim(yrbss)

yrbss %>% filter(!is.na(weight)) %>% 
ggplot(aes(x=weight))+
  geom_histogram(aes(y=..density..))+
  geom_density()+
  theme_bw()+
  labs (
    title= "Distribution of weights",
    x = "Weight (kg)",
    y = "Density"       
  )+
  NULL

```

>
The distribution of weights is heavily right skewed. Additionally, mean and median are different as well as not quite symmetricall lower and upper quartiles. 

>
We are missing 1004 observations for the variable "weight".


Next, we consider the possible relationship between a high schooler’s weight and their physical activity. Plotting the data is a useful first step because it helps us quickly visualize trends, identify strong associations, and develop research questions.

```{r}
ggplot(yrbss,aes(x=physically_active_7d,y=weight,na.rm=TRUE))+
geom_boxplot(aes(group = cut_width(physically_active_7d, 0.25)), outlier.alpha = 0.1)+
 labs (
    title= "More physicall activity has a slight positive impact on weight",
    x = "# of days of physicall activity in a week",
    y = "Weight (kg)"       
  )+
theme_bw()+
NULL


```

Let’s create a new variable `physical_3plus`, which will be `yes` if they are physically active for at least 3 days a week, and `no` otherwise.

  
```{r}
yrbss <- yrbss %>% 
  mutate(physical_3plus = ifelse(physically_active_7d >= 3, "yes", "no"))

yrbss %>% filter(!is.na(physical_3plus)) %>% 
  group_by(physical_3plus) %>% 
  summarise(count = n()) %>% 
  mutate(prop= count/sum(count))

```

Can you provide a 95% confidence interval for the population proportion of high schools that are *NOT* active 3 or more days per week?

```{r}
formula_CI <- yrbss %>% 
filter(!is.na(physical_3plus)) %>% 
  
summarise(count_no=count(physical_3plus=="no"),
            total=n(),
            proportion_no=count_no/total,
  
lower_95_CI_no=proportion_no-1.96*sqrt(
              (proportion_no*(1-proportion_no))/total),           
          
upper_95_CI_no=proportion_no+1.96*sqrt(
              (proportion_no*(1-proportion_no))/total))

formula_CI[4:5]
```
  
Let's show a boxplot of `physical_3plus` vs. `weight`. Is there a relationship between these two variables? What did you expect and why?

```{r, boxplot}
yrbss %>% 
filter(!is.na(physical_3plus)) %>% 
  ggplot(aes(x=physical_3plus,y=weight))+
  geom_boxplot(outlier.alpha = 0.1)+
  labs (
    title= "Physicall activity at least 3 times a week positively impacts weight",
    x = "'Yes' if a particular schooler is physically active at least 3 times a week and 'No' otherwise",
    y = "Weight (kg)"       
  )+
  theme_bw()+
  NULL

```

>
I expected that being physically active at least 3 times a week will result in a lower weight as one would burn more calories, but from the graph it is evident that more physicall activity is  positively correlated with weight. 

## Confidence Interval

Boxplots show how the medians of the two distributions compare, but we can also compare the means of the distributions using either a confidence interval or a hypothesis test.


```{r}
yrbss %>%
  group_by(physical_3plus) %>%
  filter(!is.na(physical_3plus)) %>% 
  summarise(mean_weight = mean(weight, na.rm = TRUE),
            sd_weight = sd(weight, na.rm=TRUE),
            count = n(),
            se_weight = sd_weight/sqrt(count),
            t_critical = qt(0.975, count-1), 
            margin_of_error = t_critical * se_weight,
            lower = mean_weight - t_critical * se_weight,
            upper = mean_weight + t_critical * se_weight
            )

```

There is an observed difference of about 1.77kg (68.44 - 66.67), and we notice that the two confidence intervals do not overlap. It seems that the difference is at least 95% statistically significant. Let us also conduct a hypothesis test.

## Hypothesis test with formula

Write the null and alternative hypotheses for testing whether mean weights are different for those who exercise at least times a week and those who don’t.

> 
Null hypotheses: true difference in mean weights for high schoolers who exercise at least 3 times a week and those who don't is equal to 0
> 
Alternative hypotheses: true difference in mean weights is not equal to 0

```{r}
t.test(weight ~ physical_3plus, data = yrbss)
```

> p-value is very close to 0. Thus, we reject the Null and confirm that true difference in mean weights is not equal to 0

## Hypothesis test with `infer`

We need to initialize the test, which we will save as `obs_diff`.

```{r}
obs_diff <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  calculate(stat = "diff in means", order = c("yes", "no"))  
  
```

After we have initialized the test, we need to simulate the test on the null distribution, which we will save as null.

```{r}
null_dist <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no"))

```

We can visualize this null distribution with the following code:

```{r}
ggplot(data = null_dist, aes(x = stat)) +
  geom_histogram()

```

Now that the test is initialized and the null distribution formed, we can visualise to see how many of these null permutations have a difference of at least `obs_stat` of `r obs_diff %>% pull() %>% round(2)`?

We can also calculate the p-value for your hypothesis test using the function `infer::get_p_value()`.

```{r}

null_dist %>% visualize() +
  shade_p_value(obs_stat = obs_diff, direction = "two-sided")

null_dist %>%
  get_p_value(obs_stat = obs_diff, direction = "two_sided")

```
> p-value is very close to 0. Thus, we reject the Null and confirm that true difference in mean weights is not equal to 0, so being physically active at least 3 times a week has an effect on weight
